\section{Background}
\label{sec:background}

\subsection{Zero-Noise Extrapolation}

Zero-noise extrapolation estimates the ideal expectation value $\langle O \rangle_0$ of an observable $O$ by measuring at multiple noise levels $\lambda_1 < \lambda_2 < \cdots < \lambda_k$ and fitting an extrapolation model $f(\lambda)$ such that $\hat{\langle O \rangle}_0 = f(0)$~\cite{Temme2017, Li2017}.

Noise amplification is achieved through unitary folding~\cite{GiurgicaTiron2020}: given a circuit $U$, the folded circuit $U(U^\dagger U)^n$ implements the same logical operation at noise scale factor $\lambda = 2n + 1$. Common extrapolation models include:
\begin{align}
    f_\text{lin}(\lambda) &= a + b\lambda, \label{eq:linear} \\
    f_\text{poly}(\lambda) &= \sum_{i=0}^{d} a_i \lambda^i, \label{eq:poly} \\
    f_\text{exp}(\lambda) &= a + b e^{-c\lambda}. \label{eq:exp}
\end{align}

The choice among these models significantly impacts accuracy: linear models underfit at high noise, while higher-order polynomials overfit with limited data points~\cite{Cai2023}.

\subsection{Physically-Bounded Extrapolation}

Miranskyy et al.~\cite{Miranskyy2026} observed that unconstrained extrapolation frequently produces unphysical estimates---values outside the domain $[l_b, u_b]$ dictated by the observable's spectrum. For probability estimation, $l_b = 0$ and $u_b = 1$; for Pauli observables, $l_b = -1$ and $u_b = 1$.

Their approach constrains the optimization:
\begin{equation}
    \hat{\theta} = \arg\min_\theta \sum_{i=1}^{k} \left[ f(\lambda_i; \theta) - y_i \right]^2 \quad \text{s.t.} \quad l_b \leq f(0; \theta) \leq u_b,
    \label{eq:bounded_opt}
\end{equation}
where $y_i = \langle O \rangle_{\lambda_i}$ are the measured values. This is solved via constrained optimization (e.g., L-BFGS-B with box constraints on the zero-noise estimate).

\subsection{Information-Theoretic Model Selection}

The Akaike Information Criterion (AIC)~\cite{Akaike1974} balances goodness-of-fit against model complexity:
\begin{equation}
    \text{AIC} = 2k - 2\ln(\hat{L}),
\end{equation}
where $k$ is the number of parameters and $\hat{L}$ is the maximized likelihood. For small samples ($n/k < 40$), the corrected version AICc~\cite{Hurvich1989} adds a bias-correction term:
\begin{equation}
    \text{AICc} = \text{AIC} + \frac{2k(k+1)}{n - k - 1},
    \label{eq:aicc}
\end{equation}
where $n$ is the number of data points. This correction is critical for ZNE, where typically $n \in \{3, 5, 7\}$ and models have $k \in \{2, 3, 4\}$ parameters---precisely the regime where AICc provides the strongest regularization against overfitting.

\subsection{Related Work}

Several approaches address ZNE limitations. Machine learning methods~\cite{Liao2024, Wang2026} train models on circuit-noise pairs but require extensive training data and lack physical guarantees. Probabilistic error cancellation (PEC)~\cite{Temme2017} provides unbiased estimates but incurs exponential sampling overhead. Physics-inspired extrapolation (PIE)~\cite{DiezValle2025} achieves constant overhead but targets a different noise regime. Proprietary solutions such as Q-CTRL Fire Opal~\cite{Mundada2023} combine multiple techniques but are closed-source and hardware-specific.

Our work is complementary: we provide physics-informed bounds with principled model selection, applicable to any ZNE implementation including Mitiq~\cite{LaRose2022}, while remaining fully open-source.
