\section{Discussion}
\label{sec:discussion}

\subsection{When Does Bounded ZNE + AICc Provide Value?}

Our results identify a precise regime where the method provides maximum benefit: circuits whose ideal expectation value lies within $\sim$5\% of a physical boundary, executed at noise levels where standard extrapolation overshoots. This regime is not a corner case---it corresponds to the practical scenario of high-fidelity circuits on improving hardware, which is precisely the trajectory of NISQ development.

As quantum hardware improves (lower $\epsilon_g$), two trends emerge: (i) ideal values become more accessible (closer to boundaries), and (ii) the relative extrapolation distance $|f(0) - y_1|/|y_1|$ increases because measurements are closer to ideal. Both trends \emph{increase} the probability of unphysical extrapolation, making our method increasingly relevant as hardware matures toward quantum advantage.

Conversely, for circuits with ideal values far from boundaries (e.g., random circuits with $\langle O \rangle \approx 0.5$) or at very high noise where all methods fail, bounded ZNE provides no advantage over standard linear ZNE. Users should apply our method selectively based on the observable being estimated.

\subsection{Computational Overhead}

The computational cost of our method is:
\begin{itemize}
    \item \textbf{Per-model fitting}: 3 L-BFGS-B optimizations, each converging in $<$50 iterations for $n \leq 7$ data points. Wall time: $<$1 second total.
    \item \textbf{AICc computation}: $O(1)$ per model (closed-form from RSS).
    \item \textbf{Total overhead vs.\ standard ZNE}: $<$3 seconds per circuit, dominated by the exponential model's nonlinear optimization. This is negligible compared to the quantum execution time (minutes to hours on real hardware).
\end{itemize}

The method does \emph{not} increase the number of quantum circuit executions---it uses the same noise-amplified measurements as standard ZNE. The only additional cost is classical post-processing.

\subsection{Scalability Considerations}

Our current benchmark covers 2--6 qubit circuits. Several factors affect scalability to larger systems:

\textbf{Noise amplification.} Unitary folding scales linearly with circuit size---folding a 100-qubit circuit is no harder than folding a 5-qubit circuit. The extrapolation and model selection are performed on scalar expectation values regardless of system size.

\textbf{Observable locality.} For local observables (e.g., single-qubit Pauli operators), the physical bounds $[-1, 1]$ apply regardless of system size. For global observables (e.g., fidelity with a target state), the bounds tighten as system size grows, potentially making our method more effective.

\textbf{Noise model validity.} At larger qubit counts, noise becomes increasingly non-Markovian and spatially correlated~\cite{vandenBerg2023}. The depolarizing noise model used in our benchmarks may not capture these effects. However, the physical bounds remain valid regardless of the noise model---they depend only on the observable's spectrum, not on the noise channel.

\textbf{Comparison with scalable alternatives.} ML-based methods~\cite{Liao2024} have demonstrated scalability to 100 qubits but require retraining for each new circuit family. PEC~\cite{vandenBerg2023} scales to 100+ qubits with sparse Pauli-Lindblad models but requires noise characterization. Our method requires neither training nor characterization, making it immediately applicable at any scale---though its advantage may diminish if noise correlations cause the true decay to deviate significantly from our candidate models.

\subsection{Limitations}

We identify several limitations that should guide future work:

\begin{enumerate}
    \item \textbf{Simulator-only validation.} All results use noise models calibrated from real hardware data, but simulator noise (depolarizing) is symmetric and Markovian. Real hardware exhibits coherent errors, crosstalk, leakage, and drift that may alter extrapolation behavior. Hardware validation is essential before deployment claims can be made.

    \item \textbf{Limited model family.} Our three candidate models (polynomial, exponential, poly-exponential) may not capture all physical noise decay patterns. Circuits with non-monotonic fidelity decay (e.g., due to coherent error interference) could defeat all three models. Expanding the model family (e.g., adding Richardson extrapolation, rational functions) is straightforward but increases the risk of overfitting that AICc must control.

    \item \textbf{Fixed noise scale factors.} We use $\lambda \in \{1, 1.5, 2, 2.5, 3\}$ uniformly. Adaptive selection of scale factors based on the observed decay~\cite{GiurgicaTiron2020} could improve both accuracy and efficiency by concentrating measurements where the signal-to-noise ratio is highest.

    \item \textbf{Single-observable estimation.} We estimate one observable at a time. For applications requiring multiple correlated observables (e.g., all Pauli terms in a Hamiltonian), joint estimation with correlated bounds could improve efficiency.

    \item \textbf{MCP server scope.} The current server supports 8 tools focused on ZNE and strategy selection. Integration with PEC, CDR, and dynamical decoupling would provide a more complete mitigation toolkit.
\end{enumerate}

\subsection{Comparison with State-of-the-Art}

\textbf{vs.\ Mitiq~\cite{LaRose2022}:} Mitiq provides ZNE infrastructure (folding, extrapolation) but delegates model selection to the user via a factory pattern. Users must manually specify \texttt{LinearFactory}, \texttt{PolyFactory}, or \texttt{ExpFactory}---precisely the choice our AICc criterion automates. Our method is designed as a composable extension: replacing Mitiq's factory with an AICc-selecting factory requires $<$20 lines of code, preserving full compatibility with Mitiq's folding and benchmarking infrastructure.

\textbf{vs.\ Q-CTRL Fire Opal~\cite{Mundada2023}:} Fire Opal operates at a fundamentally different level: it combines (i) optimized control pulses that suppress errors at the physical layer, (ii) dynamical decoupling sequences, and (iii) proprietary error mitigation post-processing. Its reported $>$1000$\times$ improvement~\cite{Mundada2023} reflects the \emph{combined} effect of all three layers, making direct comparison with any single mitigation technique (including ours) methodologically inappropriate.

Key differences: Fire Opal requires IBM Quantum hardware access and is closed-source---its internal algorithms cannot be inspected, reproduced, or extended by the community. It does not publish per-technique ablation studies, making it impossible to isolate the contribution of its mitigation layer alone. Our approach is (i) fully open-source, (ii) hardware-agnostic, (iii) focused specifically on the extrapolation step with provable guarantees, and (iv) composable with other techniques including pulse optimization. The two approaches are complementary rather than competing: one could apply Fire Opal's error suppression \emph{and} our bounded ZNE for the extrapolation step.

\textbf{vs.\ PEC~\cite{Temme2017, vandenBerg2023}:} Probabilistic error cancellation provides \emph{unbiased} estimates---a stronger guarantee than our bounded estimates, which may be biased when the constraint is active. However, PEC's sampling overhead scales as $O(e^{2\gamma})$ where $\gamma = \sum_i \gamma_i$ is the total noise strength, making it impractical for deep circuits. Van den Berg et al.~\cite{vandenBerg2023} demonstrated PEC on 100-qubit circuits using sparse Pauli-Lindblad noise models, but required extensive noise characterization ($O(n^2)$ Pauli fidelity measurements). Our method requires \emph{zero} noise characterization---only the standard ZNE measurements.

\textbf{vs.\ CDR and vnCDR~\cite{Czarnik2021, Lowe2021}:} Clifford Data Regression learns a linear correction from near-Clifford training circuits that can be efficiently simulated classically. This provides excellent accuracy when the noise is well-approximated by a global depolarizing channel, but degrades under spatially correlated or non-Markovian noise. The training cost ($O(2^n)$ classical simulation per training circuit) limits scalability. Variable-noise CDR (vnCDR)~\cite{Lowe2021} combines CDR with ZNE, using multiple noise levels of training data. Our method avoids training entirely, at the cost of weaker accuracy guarantees in regimes where CDR's linear correction is well-calibrated.

\textbf{vs.\ ML-QEM~\cite{Liao2024, Wang2026, Strikis2021}:} Machine learning approaches represent the current accuracy frontier: Liao et al.~\cite{Liao2024} demonstrated ML-based mitigation on 100-qubit IBM circuits, and Wang et al.~\cite{Wang2026} achieved zero-shot transfer across circuit families using graph neural networks. These methods achieve superior accuracy but require (i) $O(10^4)$ training circuits per noise regime, (ii) retraining when hardware drifts, and (iii) provide no physical guarantees on outputs. Our method occupies a different point in the accuracy-cost-guarantee tradeoff space: lower accuracy ceiling but zero training cost and provable physical validity.

\subsection{Broader Impact: AI-Accessible Quantum Computing}

The MCP server component addresses a growing need in scientific computing: as large language models increasingly assist researchers~\cite{Kim2023}, they require access to domain-specific tools that encode expert knowledge while providing safety guardrails.

Our server demonstrates three principles for AI-accessible scientific tools:
\begin{enumerate}
    \item \textbf{Physics-informed guardrails:} The bounded ZNE tool cannot produce unphysical results regardless of how the AI agent invokes it---the physical constraints are embedded in the tool, not delegated to the agent.
    \item \textbf{Intelligent defaults:} AutoMitigator selects strategies based on circuit properties, so agents need not understand noise physics to make reasonable mitigation choices.
    \item \textbf{Transparency:} Every tool returns not just results but also metadata (model selected, AICc scores, confidence) enabling agents to reason about reliability.
\end{enumerate}

This ``AI-in-the-loop'' paradigm differs from both fully manual mitigation (requiring expert knowledge) and fully automated ML approaches (requiring training data). It provides guardrails (physical bounds) while leveraging AI agents' ability to orchestrate complex multi-step workflows---characterize noise, select strategy, execute mitigation, validate results---without human intervention at each step.
