\section{Introduction}
\label{sec:introduction}

The noisy intermediate-scale quantum (NISQ) era~\cite{Preskill2018} demands error mitigation techniques that are simultaneously reliable, automated, and physically consistent. Among available approaches, zero-noise extrapolation (ZNE)~\cite{Temme2017, Li2017} has emerged as the most widely deployed technique due to its simplicity: amplify noise systematically, measure at multiple noise levels, and extrapolate to the zero-noise limit. ZNE requires no knowledge of the noise channel, no ancilla qubits, and imposes only a polynomial sampling overhead---properties that have made it the default mitigation strategy in both Qiskit~\cite{Qiskit2024} and Mitiq~\cite{LaRose2022}.

Yet ZNE harbors two fundamental failure modes that undermine its reliability in practice:

\textbf{Failure mode 1: Unphysical predictions.} Standard polynomial or exponential extrapolation models are unconstrained---they can produce estimates outside the physical domain of the observable. For probability estimation, this means predictions $\hat{p} > 1$ or $\hat{p} < 0$; for Pauli observables, $|\hat{\langle O \rangle}| > 1$. Miranskyy et al.~\cite{Miranskyy2026} demonstrated this failure across 180,000 circuits: unconstrained linear extrapolation produces unphysical results in up to 12\% of configurations at low noise. These failures are not merely aesthetic---they propagate through downstream calculations (e.g., energy estimation in VQE) and can cause optimization algorithms to diverge.

\textbf{Failure mode 2: Unprincipled model selection.} ZNE practitioners must choose an extrapolation model (linear, quadratic, exponential, Richardson) \emph{a priori}, typically based on intuition or convention. This choice significantly impacts accuracy~\cite{Cai2023}, yet no principled selection criterion exists for the small-sample regime inherent to ZNE. With only $n \in \{3, 5, 7\}$ data points and models with $k \in \{2, 3, 4\}$ parameters, standard cross-validation is inapplicable, and residual-based selection systematically favors overfitting.

\textbf{The gap.} Existing solutions address these failures in isolation. Physically-bounded ZNE~\cite{Miranskyy2026} eliminates unphysical predictions but still requires manual model selection. Machine learning approaches~\cite{Liao2024, Wang2026, Strikis2021} learn model selection from data but require extensive training sets, lack physical guarantees, and do not generalize across noise regimes. Probabilistic error cancellation (PEC)~\cite{Temme2017, vandenBerg2023} provides unbiased estimates but incurs exponential sampling overhead. No existing method simultaneously guarantees physical validity \emph{and} provides principled, automatic model selection with zero training overhead.

\textbf{Our contribution.} We close this gap by combining two well-established tools from statistics and optimization:
\begin{enumerate}
    \item \textbf{Constrained optimization} with L-BFGS-B enforcing spectral bounds on the zero-noise estimate, guaranteeing $l_b \leq f(0;\hat\theta) \leq u_b$ for any fitted model.
    \item \textbf{AICc model selection}~\cite{Akaike1974, Hurvich1989, Burnham2002} providing a bias-corrected, information-theoretic criterion that penalizes model complexity proportional to $2k(k+1)/(n-k-1)$---a correction that becomes dominant precisely in the ZNE regime where $n/k < 10$.
\end{enumerate}

The combination is more than the sum of its parts: physical bounds prevent catastrophic extrapolation failures, while AICc prevents the subtler failure of overfitting that bounds alone cannot address (a bounded but overfitted model still produces poor estimates within $[l_b, u_b]$).

We validate this approach across 125 circuit-noise configurations using noise models calibrated from real Origin Wukong 180 hardware (169 qubits, 396 CZ gates), demonstrating:
\begin{itemize}
    \item Complete elimination of unphysical predictions (0\% vs.\ 4\% for linear ZNE)
    \item 2\% lower MAE at low noise ($\epsilon_\text{CZ} = 1\%$), the operationally relevant regime
    \item Automatic model selection requiring zero hyperparameter tuning
    \item Negligible computational overhead ($<$3 seconds per circuit)
\end{itemize}

Additionally, we release the first Model Context Protocol (MCP)~\cite{Anthropic2024} server for quantum error mitigation, verified with production AI agents (OpenAI Codex, GitHub Copilot), demonstrating that physics-informed mitigation can be made accessible to non-expert users through natural language interaction.

The remainder of this paper is organized as follows. \Cref{sec:background} reviews ZNE, model selection theory, and related work. \Cref{sec:methods} presents our algorithm with theoretical justification. \Cref{sec:results} reports comprehensive benchmarks. \Cref{sec:discussion} provides detailed analysis of when and why our method succeeds, its limitations, and comparison with state-of-the-art alternatives. \Cref{sec:conclusion} concludes with future directions.
